---
  title: "CAR Model, Mayer"
author: "Claire Kelling"
date: "June 11, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this code is to fit the CAR model to voting data, aggregated by ward. We only have Census data by district and/or block group so this presents a spatial mismatch file. 

First, we will install all necessary packages.
```{r}
# Packages:
library(sp)
library(spdep)
library(classInt)
library(fields)
library(ggplot2)
library(dplyr)
library(ade4) 
library(igraph) 
library(CARBayesdata)
library(CARBayes)
library(gridExtra)
library(xtable)
library(stringi)
require(rgdal)
library(ngspatial)
library(plyr)

```

Now, we will load the data, including the following:
  * shape files
* demographics
* voting data
```{r}
#ward shape files
#wi_ward
ward_shp <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data", layer = "wiward")

#voting and demographic data
#quick fit with demographics
demog_vote <- read.csv("C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/QuickFitWithDemographics.csv")

```

Exploring shape file
```{r}
plot(ward_shp)
```


Now we need to join the spatial polygons and the demographic/voting data.
```{r}
#re-formatting to add the data to the 
View(ward_shp@data)
ward_shp@data <- ward_shp@data[,-c(14:ncol(ward_shp@data))]

#check classes
class(ward_shp@data$OBJECTID_1)

#change column name for ID variables for which we will join
colnames(demog_vote)[2] <- "GEOID"
colnames(ward_shp@data)[1] <- "GEOID"

#change to character variable
demog_vote$GEOID <- as.character(demog_vote$GEOID)
ward_shp@data$GEOID <- as.character(ward_shp@data$GEOID)

#creating new variable
demog_vote$IncumD_Pop <- demog_vote$Dincumbent*demog_vote$PERSONS18
demog_vote$IncumR_Pop <- demog_vote$Rincumbent*demog_vote$PERSONS18

#join on GEOID
ward_shp@data <- left_join(ward_shp@data, demog_vote, by = (GEOID = "GEOID"))

length(which(is.na(ward_shp@data$ED01)))
nrow(ward_shp@data)

nrow(demog_vote)
nrow(ward_shp@data)

class(demog_vote$GEOID)
class(ward_shp@data$GEOID)

#BBPRESDEM12 + BBPRESREP12+PERSONS18 +BLACK18 + WHITE18 + HISPANIC18+ IncumD_Pop + IncumR_Pop

class(ward_shp@data$BBWSADEM12) 
class(ward_shp@data$BBWSADEM12) 
class(ward_shp@data$BBPRESDEM12) 
class(ward_shp@data$BBPRESREP12)


ward_shp@data$BBWSADEM12 <- round(as.numeric(ward_shp@data$BBWSADEM12))
ward_shp@data$BBWSADEM12 <- round(as.numeric(ward_shp@data$BBWSADEM12))
ward_shp@data$BBPRESDEM12 <- round(as.numeric(ward_shp@data$BBPRESDEM12))
ward_shp@data$BBPRESREP12 <- round(as.numeric(ward_shp@data$BBPRESREP12))
ward_shp@data$CWSADEM12 <- round(ward_shp@data$CWSADEM12)
ward_shp@data$CWSAREP12 <- round(ward_shp@data$CWSAREP12)

#we have successfully joined the two datasets
```


Our first step in terms of modeling will be to run our model in a linear model and to test for spatial dependence. 
```{r}
#we need to decide variables to use in our analysis
colnames(demog_vote)
#contested - B if it is contested
#Mayer: BBPRESDEM12, BBPRESREP12, PERSONS18, BLACK18, HISPANIC18
#Mayer Response: CWSADEM12, CWSAREP12
#TP (third party)

#approximately 50% of the data is contested (vs uncontested)
#right now, I will just use the uncontested data
#length(c(which(ward_shp$CWSADEM12 == 0 ), which(ward_shp$CWSAREP12==0)))
#unc <- unique(c(which(ward_shp$CWSADEM12 == 0 ), which(ward_shp$CWSAREP12 == 0 )))

#length(which(ward_shp$CWSADEM12 == 0  & ward_shp$CWSAREP12 == 0 ))

unc <- unique(which(ward_shp@data$Contested == "U"))
con_shp <- ward_shp[-unc,]

#need to delete Object ID 6481 (outlier)


#non-spatial modeling
d.form <- CWSADEM12 ~ CPREDEM12 + CPREREP12+PERSONS18 +BLACK18 + WHITE18 + HISPANIC18+ IncumD_Pop + IncumR_Pop #+ County#try adding fixed effects for county
r.form <- CWSAREP12 ~ CPREDEM12 + CPREREP12+PERSONS18 +BLACK18 + WHITE18 + HISPANIC18+ IncumD_Pop + IncumR_Pop #+ County#try adding fixed effects for county


c.d.model <- lm(formula=d.form, data=con_shp@data)
c.d.resid.model <- con_shp@data$CWSADEM12 - c.d.model$fitted.values
c.r.model <- lm(formula=r.form, data=con_shp@data)
c.r.resid.model <- con_shp@data$CWSAREP12 - c.r.model$fitted.values



#glm model
c.d.glm_model <- glm(formula=d.form, family = "poisson", data=con_shp@data)
c.d.resid.glm <- con_shp@data$CWSADEM12 - c.d.glm_model$fitted.values
c.r.glm_model <- glm(formula=r.form, family = "poisson", data=con_shp@data)
c.r.resid.glm <- con_shp@data$CWSAREP12 - c.r.glm_model$fitted.values


#test for spatial dependence
#null hypothesis of no spatial autocorrelation (alternative of positive spatial autocorrelation)
#also computes Moran's I statistic 
#if p-value < 0.05, we conclude there is positve spatial autocorrelation
W.nb.con <- poly2nb(con_shp, row.names = rownames(con_shp@data))

##############
### Creating adjacency matrix from nb
##############
W_mat_con <- nb2mat(W.nb.con, style="B", zero.policy=TRUE)


#############
### Moran's I Test
#############

W.list.con <- nb2listw(W.nb.con, style="B", zero.policy = TRUE)

moran.mc(x=c.d.resid.glm, listw=W.list.con, nsim=1000, zero.policy = TRUE)
moran.mc(x=c.r.resid.glm, listw=W.list.con, nsim=1000, zero.policy = TRUE)

#if p-value < 0.05, we conclude there is positve spatial autocorrelation

```


Need to eliminate the islands with no neighbors
```{r}
no_nb <- which(rowSums(W_mat_con) == 0) #13 with no neighbors


nonb_W <- W_mat_con[-no_nb, -no_nb]
nonb_shp <- con_shp[-no_nb,]

nonb_Wnb <- poly2nb(nonb_shp, row.names = rownames(nonb_shp@data))
```


Now, we will fit both the Leroux and the BYM model.
```{r}
##############
#Leroux Model
##############
#d.model.ler <- S.CARleroux(formula=d.form, data=no_dem@data,
#                             family="poisson", W=W_mat_nod, burnin=20000, n.sample=120000, thin=10)

#r.model.ler <- S.CARleroux(formula=r.form, data=no_rep@data,
#                             family="poisson", W=W_mat_nor, burnin=20000, n.sample=120000, thin=10)

#started at 3:07pm, finished in  2162.1 seconds
c.d.model.ler <- S.CARleroux(formula=d.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)
#started at 3:50pm, started MCMC at 3;58pm
c.r.model.ler <- S.CARleroux(formula=r.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)

##############
# BYM Model
##############
#d.model.bym <- S.CARbym(formula=d.form, data=no_dem@data,
#                          family="poisson", W=W_mat_nod, burnin=20000, n.sample=120000, thin=10)

#r.model.bym <- S.CARbym(formula=r.form, data=no_rep@data,
#                          family="poisson", W=W_mat_nor, burnin=20000, n.sample=120000, thin=10)

c.d.model.bym <- S.CARbym(formula=d.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)


c.r.model.bym <- S.CARbym(formula=r.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)


#save(c.d.model.ler, #c.r.model.ler, c.d.model.bym, c.r.model.bym, 
     file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/prelim_fit/c_d_model_ler.Rdata")
save(c.r.model.ler, #c.r.model.ler, c.d.model.bym, c.r.model.bym, 
     file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/prelim_fit/c_r_model_ler.Rdata")

#rm(c.d.model.ler)
rm(c.r.model.ler)

##############
# Model Evaluation
##############
#  dem ler
#d.model.ler$modelfit
#  rep ler
#d.model.ler$modelfit
#  dem bym
#r.model.bym$modelfit
#  rep bym
#r.model.bym$modelfit

#Only using contested precincts:
#  dem ler
c.d.model.ler$modelfit
#  rep ler
c.r.model.ler$modelfit
#  dem bym
c.d.model.bym$modelfit
#  rep bym
c.r.model.bym$modelfit
```

Now, I will also include a model that accounts for spatial autocorrelation.
```{r}
#d.sp.sglmm.fit <- sparse.sglmm(formula = d.form,data=no_dem@data, family = poisson, A = W_mat_nod,
#                             verbose = TRUE) #tune = list(sigma.s = 0.02)

#r.sp.sglmm.fit <- sparse.sglmm(formula = r.form,data=no_rep@data, family = poisson, A = W_mat_nor,
#                             verbose = TRUE) #tune = list(sigma.s = 0.02)

rownames(nonb_W) <- NULL
c.d.sp.sglmm.fit <- sparse.sglmm(formula = d.form,data=nonb_shp@data, family = poisson, 
                                 A = nonb_W, verbose = TRUE) #tune = list(sigma.s = 0.02)

#25 min at 9:43, stuck at 44%, 23m 49s

c.r.sp.sglmm.fit <- sparse.sglmm(formula = r.form,data=nonb_shp@data, family = poisson, 
                                 A = nonb_W, verbose = TRUE) #tune = list(sigma.s = 0.02)
#11%, 32m 22s remaining at 9:56pm
#55% ~19m 27s  

#save the models for later reference, takes over an hour to run
save(c.d.model.ler, c.r.model.ler, c.d.model.bym, c.r.model.bym, 
     file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/prelim_fit/prelim_mod_fit2.Rdata")

#summary(sp.sglmm.fit)
#d.sp.sglmm.fit$dic
#d.sp.sglmm.fit$pD

#r.sp.sglmm.fit$dic
#r.sp.sglmm.fit$pD

c.d.sp.sglmm.fit$dic
c.d.sp.sglmm.fit$pD

c.r.sp.sglmm.fit$dic
c.r.sp.sglmm.fit$pD

```

Now, we need to compare coefficients to see if there was any spatial autocorrelation.
```{r}
model.bym$summary.results[,1:3]

summary(c.d.sp.sglmm.fit)
```

#Model Evaluation
Now, I will compare the models using mean-squared prediction error.

First, I need to set up predicted for each model and actual for both Democratic and Republican races.
```{r}
actual.d <- nonb_shp@data$CWSADEM12
actual.r <- nonb_shp@data$CWSAREP12

d.pred.bym <- c.d.model.bym$fitted.values
r.pred.bym <- c.r.model.bym$fitted.values

d.pred.ler <- c.d.model.ler$fitted.values
r.pred.ler <- c.r.model.ler$fitted.values

#d.pred.sglmm <- c.d.sp.sglmm.fit$fitted.values
#r.pred.sglmm <- c.r.sp.sglmm.fit$fitted.values

# actual_tab <- cbind(nonb_shp@data$District, actual.d, actual.r)
# actual_tab <- as.data.frame(actual_tab)
# colnames(actual_tab) <- c("District", "actual_d", "actual_r")
# test <- actual_tab %>% group_by(District) %>% dplyr::summarise(dem_votes=sum(actual_d))

actual_tab <- cbind(con_shp@data$District, con_shp@data$CWSADEM12, con_shp@data$CWSAREP12)
actual_tab <- as.data.frame(actual_tab)
colnames(actual_tab) <- c("District", "actual_d", "actual_r")
actual_tab2 <- actual_tab %>% group_by(District) %>% dplyr::summarise(dem_votes=sum(actual_d), rep_votes = sum(actual_r))

length(which(actual_tab2[,2] > actual_tab2[,3])) #6 times that the actual_tab was greater (Samir said 5?)

```

Now, I will compute the MSPE for each model.
```{r}
#### Democratic Prediction:
#linear model
d.lm.mspe <- mean((c.d.resid.model)^2)
#glm
d.glm.mspe <- mean((c.d.resid.glm)^2)
#bym
d.bym.mspe <- mean((d.pred.bym-actual.d)^2)
#leroux
d.ler.mspe <- mean((d.pred.ler-actual.d)^2)
#sglmm
#d.sglmm.mspe <- mean((d.pred.sglmm - actual.d)^2)

#### Republican Prediction:
#linear model
r.lm.mspe <- mean((c.r.resid.model)^2)
#glm
r.glm.mspe <- mean((c.r.resid.glm)^2)
#bym
r.bym.mspe <- mean((r.pred.bym-actual.r)^2)
#leroux
r.ler.mspe <- mean((r.pred.ler-actual.r)^2)
#sglmm
#r.sglmm.mspe <- mean((r.pred.sglmm - actual.r)^2)

mspe_tab <- rbind(c(d.lm.mspe, d.glm.mspe, d.bym.mspe, d.ler.mspe),#, d.sglmm.mspe),
                  c(r.lm.mspe, r.glm.mspe, r.bym.mspe, r.ler.mspe))#, r.sglmm.mspe))
#mspe_tab_count <- mspe_tab
colnames(mspe_tab) <- c("lin", "glm", "bym", "leroux")#, "sglmm")
rownames(mspe_tab) <- c("Democrat", "Republican")

```


Bayesian Model Comparison Techniques:
```{r}
#smaller DIC is better
#higher perc dev expl is better
#leroux models are much better and explain much of the deviance (than bym)
# dem ler
c.d.model.ler$modelfit
#  rep ler
c.r.model.ler$modelfit
#  dem bym
c.d.model.bym$modelfit
#  rep bym
c.r.model.bym$modelfit

bayes_mod_eval <- rbind(c(c.d.model.bym$modelfit[1], c.d.model.ler$modelfit[1], 
                          c.r.model.bym$modelfit[1], c.r.model.ler$modelfit[1]),
                        c(c.d.model.bym$modelfit[7], c.d.model.ler$modelfit[7], 
                          c.r.model.bym$modelfit[7], c.r.model.ler$modelfit[7]))
bayes_mod_eval <- as.data.frame(bayes_mod_eval)
colnames(bayes_mod_eval) <- c("BYM Dem", "Ler Dem", "BYM Rep", "Ler Rep")
rownames(bayes_mod_eval) <- c("DIC", "perc dev explained")

#c.d.sp.sglmm.fit$dic
#c.d.sp.sglmm.fit$pD

#c.r.sp.sglmm.fit$dic
#c.r.sp.sglmm.fit$pD
```


Creation of counterfactual scenario where there is no incumbent:
```{r}
#First, I will create the model dataframe
mod_dat <- demog_vote[,c("BBPRESDEM12", "BBPRESREP12", "PERSONS18", "BLACK18", "WHITE18", "HISPANIC18", "IncumD_Pop", "IncumR_Pop")]
mod_dat$int <- rep(1, nrow(mod_dat))
mod_dat <- cbind(mod_dat[,9], mod_dat[,1:8])

mod_dat2 <- con_shp@data[,c("BBPRESDEM12", "BBPRESREP12", "PERSONS18", "BLACK18", "WHITE18", "HISPANIC18", "IncumD_Pop", "IncumR_Pop")]
mod_dat2$int <- rep(1, nrow(mod_dat2))
mod_dat2 <- cbind(mod_dat2[,9], mod_dat2[,1:8])

#create function that returns predicted values under counterfactual situation
pred_val <- function(coeffs.Dem, coeffs.Rep, mod_dat){
  #mod_dat <- mod_dat2
  coeffs.Dem[8] <- 0
  coeffs.Dem[9] <- 0
  coeffs.Rep[8] <- 0
  coeffs.Rep[9] <- 0
  
  tableDem <- as.matrix(mod_dat)%*%coeffs.Dem
  tableDem <- as.numeric(tableDem)
  tableDem <- exp(tableDem) #comment out when not glm
  tableDem <- cbind(tableDem, demog_vote$District)
  #tableDem <- cbind(tableDem, con_shp@data$District)
  tableDem <- as.data.frame(tableDem)
  colnames(tableDem) <- c("votes", "District")
  
  length(unique(tableDem$District))
  
  v_table_d <- tableDem %>% group_by(District) %>% dplyr::summarise(dem_votes=sum(votes))
  
  tableRep <- as.matrix(mod_dat)%*%coeffs.Rep
  tableRep <- as.numeric(tableRep)
  tableRep <- exp(tableRep) #comment out when not glm
  tableRep <- cbind(tableRep, demog_vote$District)
  #tableRep <- cbind(tableRep, con_shp@data$District)
  tableRep <- as.data.frame(tableRep)
  colnames(tableRep) <- c("votes", "District")
  
  v_table_r <- tableRep %>% group_by(District) %>% dplyr::summarise(rep_votes=sum(votes))
  
  v_table <- left_join(v_table_d, v_table_r)
  
  colnames(v_table) <- c("District", "Dem. Votes", "Rep. Votes")
  v_table2 <- v_table
  return(v_table)
}


######
#Linear Model:
######

coeffs.Dem <-summary(c.d.model)$coefficients[,1]
coeffs.Rep <-summary(c.r.model)$coefficients[,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_lm <- pred_val(coeffs.Dem, coeffs.Rep, mod_dat)
View(cbind(v_table_lm, actual_tab2))

######
#GLM:
######

#test3 <- coef(c.d.glm_model)
#View(cbind(coeffs.Dem, test3))

coeffs.Dem <-summary(c.d.glm_model)$coefficients[,1]
coeffs.Rep <-summary(c.r.glm_model)$coefficients[,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_glm <- pred_val(coeffs.Dem, coeffs.Rep, mod_dat)
View(cbind(v_table_glm, actual_tab2))

length(which(v_table_glm[,2] > v_table_glm[,3])) #still only predicts 3


######
#BYM:
######

coeffs.Dem <- c.d.model.bym$summary.results[1:9,1]
coeffs.Rep <- c.r.model.bym$summary.results[1:9,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_bym <- pred_val(coeffs.Dem, coeffs.Rep, mod_dat)
length(which(v_table_bym[,2] > v_table_bym[,3])) #still only predicts 3

######
#Leroux:
######

coeffs.Dem <- c.d.model.ler$summary.results[1:9,1]
coeffs.Rep <- c.r.model.ler$summary.results[1:9,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_ler <- pred_val(coeffs.Dem, coeffs.Rep, mod_dat)
length(which(v_table_ler[,2] > v_table_ler[,3])) #still only predicts 2

######
#SGLMM:
######

# coeffs.Dem <- c.d.sp.sglmm.fit$coefficients
# coeffs.Rep <- c.r.sp.sglmm.fit$coefficients
# 
# coeffs.Dem <- as.numeric(coeffs.Dem)
# coeffs.Rep <- as.numeric(coeffs.Rep)
# 
# v_table_sglmm <- pred_val(coeffs.Dem, coeffs.Rep)

write.csv(v_table_lm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/lin_mod.csv')
write.csv(v_table_glm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/glm.csv')
write.csv(v_table_bym, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/bym.csv')
write.csv(v_table_ler, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/ler.csv')
#write.csv(v_table_sglmm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/sglmm.csv')```
```

Now, we will consider county fixed effects in our counterfactual scenario.

```{r}

#First, I will create the model dataframe
mod_dat <- demog_vote[,c("BBPRESDEM12", "BBPRESREP12", "PERSONS18", "BLACK18", "WHITE18", "HISPANIC18", "IncumD_Pop", "IncumR_Pop")]
d_mod_dat <- con_shp@data[,c("CWSADEM12", "BBPRESDEM12", "BBPRESREP12", "PERSONS18", "BLACK18", "WHITE18", "HISPANIC18", "IncumD_Pop", "IncumR_Pop")]
r_mod_dat <- con_shp@data[,c("CWSAREP12", "BBPRESDEM12", "BBPRESREP12", "PERSONS18", "BLACK18", "WHITE18", "HISPANIC18", "IncumD_Pop", "IncumR_Pop")]


mod_mat <- model.matrix( ~ County - 1, data=con_shp@data )
d_mod_dat <- cbind(d_mod_dat, mod_mat)
r_mod_dat <- cbind(r_mod_dat, mod_mat)
mod_mat2 <- model.matrix( ~ County - 1, data=demog_vote)
mod_dat$int <- rep(1, nrow(mod_dat))
mod_dat <- cbind(mod_dat[,9], mod_dat[,1:8], mod_mat2)
colnames(mod_dat)[1]<- c("int")

#non-spatial modeling
d.form <- CWSADEM12 ~ .
r.form <- CWSAREP12 ~ .

c.d.model <- lm(formula=d.form, data=d_mod_dat)
c.d.resid.model <- con_shp@data$CWSADEM12 - c.d.model$fitted.values
c.r.model <- lm(formula=r.form, data=r_mod_dat)
c.r.resid.model <- con_shp@data$CWSAREP12 - c.r.model$fitted.values

#glm model

c.d.glm_model <- glm(formula=d.form, family = "poisson", data=d_mod_dat)
c.d.resid.glm <- con_shp@data$CWSADEM12 - c.d.glm_model$fitted.values
c.r.glm_model <- glm(formula=r.form, family = "poisson", data=r_mod_dat)
c.r.resid.glm <- con_shp@data$CWSAREP12 - c.r.glm_model$fitted.values

######
#Linear Model:
######

#take out columns with column sum = 0
test <- mod_dat[,-as.numeric(which(colSums(d_mod_dat) == 0))]
test <- test[,-32] #also taking out CountyYukon-Koyukuk Census Area`

coeffs.Dem <-summary(c.d.model)$coefficients[,1]
coeffs.Rep <-summary(c.r.model)$coefficients[,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

co_v_table_lm <- pred_val(coeffs.Dem, coeffs.Rep, test)
length(which(co_v_table_lm[,2] > co_v_table_lm[,3])) #still only predicts 4

######
#GLM:
######

coeffs.Dem <-summary(c.d.glm_model)$coefficients[,1]
coeffs.Rep <-summary(c.r.glm_model)$coefficients[,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

co_v_table_glm <- pred_val(coeffs.Dem, coeffs.Rep, test)

length(which(co_v_table_glm[,2] > co_v_table_glm[,3])) #still only predicts 4


#linear model
co.d.lm.mspe <- mean((c.d.resid.model)^2)
co.r.lm.mspe <- mean((c.r.resid.model)^2)
#glm
co.d.glm.mspe <- mean((c.d.resid.glm)^2)
co.r.glm.mspe <- mean((c.r.resid.glm)^2)

mspe_tab <- rbind(c(d.lm.mspe, d.glm.mspe, d.bym.mspe, d.ler.mspe, co.d.lm.mspe, co.d.glm.mspe),#, d.sglmm.mspe),
                  c(r.lm.mspe, r.glm.mspe, r.bym.mspe, r.ler.mspe, co.r.lm.mspe, co.r.glm.mspe))#, r.sglmm.mspe))
#mspe_tab_count <- mspe_tab
colnames(mspe_tab) <- c("lin", "glm", "bym", "leroux", "county lm", "county glm")#, "sglmm")
rownames(mspe_tab) <- c("Democrat", "Republican")
View(mspe_tab)

write.csv(co_v_table_lm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/co_lin_mod.csv')
write.csv(co_v_table_glm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/co_glm.csv')

#save workspace
save.image(file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Mayer Project/data/mod_fit/final_workspace.Rdata")

```


To Do: 
* need to also incorporate incumbency
* cross validation statistics?
* try including and not the county indicator
* hypotheticaly if there were no incumbents, fit the model on precincts and then aggregate to the house district level, and see who would win
** in other words, we want two columns (repub votes and dem votes) for each of the 40 districts - do a group_by
* Histogram EG and # D Seats
*Also mark boundary lines of districts