---
  title: "CAR Model, Alaska"
author: "Claire Kelling"
date: "June 11, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this code is to fit the CAR model to voting data, aggregated by ward. We only have Census data by district and/or block group so this presents a spatial mismatch file. 

First, we will install all necessary packages.
```{r}
# Packages:
library(sp)
library(spdep)
library(classInt)
library(fields)
library(ggplot2)
library(dplyr)
library(ade4) 
library(igraph) 
library(CARBayesdata)
library(CARBayes)
library(gridExtra)
library(xtable)
library(stringi)
require(rgdal)

```

Now, we will load the data, including the following:
  * shape files
* demographics
* voting data
```{r}
#shape files
load(file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/precinct_shape.Rdata")

#initial precinct shape files, showing some wrong shapes
prec_shp <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2013-SW-Precinct-Proc-Plan", layer = "2013-SW-Proc-Shape-files")

dist_shp <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2013-HD-ProclamationPlan", layer = "2013ProclamationPlan")

#voting and demographic data
demog_vote <- read.csv("C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/AKDataByPrecinct.csv")

#redo shape file from github source
# https://github.com/aaron-strauss/precinct-shapefiles/blob/master/ak/2012_ak_precincts.zip
prec_shp2 <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2012_ak_precincts", layer = "SW Amended Precinct shape files")

```

Exploring shape file
```{r}
plot(prec_shp)
plot(dist_shp)

#new shape file for precincts
plot(prec_shp2) #still has the precincts on the right hand side of the graph
```


Now we need to join the spatial polygons and the demographic/voting data.
```{r}
#re-formatting to add the data to the 
View(prec_shp@data)

#check classes
class(prec_shp@data$DISTRICT)

#change column name for ID variables for which we will join
colnames(demog_vote)[1] <- "GEOID"
colnames(prec_shp@data)[3] <- "GEOID"

#change to character variable
demog_vote$GEOID <- as.character(demog_vote$GEOID)
prec_shp@data$GEOID <- as.character(prec_shp@data$GEOID)

#join on GEOID
prec_shp@data <- left_join(prec_shp@data, demog_vote, by = (GEOID = "GEOID"))

length(which(is.na(test$sumRENTED)))
nrow(test)

nrow(demog_vote)
nrow(prec_shp@data)

class(demog_vote$GEOID)
class(prec_shp@data$GEOID)
```


Our first step in terms of modeling will be to run our model in a linear model and to test for spatial dependence. 
```{r}
#we need to decide variables to use in our analysis
colnames(demog_vote)
#contested - B if it is contested
#Mayer: D.Pres.Vote, R.Pres.Vote, sumTOTALPOP, sumBLACK, sumHISPANIC
#Mayer Response: D.Dist.Vote, R.Dist.Vote
#TP (third party)

class(prec_shp@data$R.Pres.Vote)

#non-spatial modeling
form <- D.Dist.Vote ~ D.Pres.Vote + R.Pres.Vote+sumTOTALPOP +sumBLACK + sumHISPANIC
model <- lm(formula=form, data=prec_shp@data)
resid.model <- residuals(model)

#glm model
glm_model <- glm(formula=form, family = "poisson", data=prec_shp@data)


#test for spatial dependence
#null hypothesis of no spatial autocorrelation (alternative of positive spatial autocorrelation)
#also computes Moran's I statistic 
#if p-value < 0.05, we conclude there is positve spatial autocorrelation
W.nb <- poly2nb(prec_shp, row.names = rownames(prec_shp@data))

coords <- coordinates(prec_shp)
plot(W.nb, coords, pch = 1, cex = 0.6, add = TRUE)


##############
### Creating adjacency matrix from nb
##############
W_mat <- nb2mat(W.nb, style="B")

#############
### Moran's I Test
#############

#addition
W.add.list <- nb2listw(add_nb, style="B")
moran.mc(x=resid.model, listw=W.add.list, nsim=1000)

```


Now, we will fit both the Leroux and the BYM model.
```{r}
##############
#Leroux Model
##############
#binary
model.ler.bin <- S.CARleroux(formula=form, data=precinct_shape@data,
                             family="poisson", W=bin_W, burnin=20000, n.sample=120000, thin=10)

##############
# BYM Model
##############
#binary
model.bym.bin <- S.CARbym(formula=form, data=precinct_shape@data,
                          family="poisson", W=add_W, burnin=20000, n.sample=120000, thin=10)

#save the models for later reference, takes over an hour to run
save(model.bym.add, model.bym.bin, model.ler.add, model.ler.bin, file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/prelim_fit/precinct_shape.Rdata")

##############
# BYM Model
##############
#  add bym
model.bym.add$modelfit
#  bin bym
model.bym.bin$modelfit
#  add ler
model.ler.add$modelfit
#  bin ler
model.ler.bin$modelfit
```

Now, I will also include a model that accounts for spatial autocorrelation.
```{r}
sp.sglmm.fit <- sparse.sglmm(formula = form,data=precinct_shape_geog@data, family = poisson, A = W_mat,
                             verbose = TRUE) #tune = list(sigma.s = 0.02)


#summary(sp.sglmm.fit)
sp.sglmm.fit$dic
sp.sglmm.fit$pD
```

Now, we need to compare coefficients to see if there was any spatial autocorrelation.
```{r}
model.bym.geog$summary.results[,1:3]

summary(sp.sglmm.fit.soc)
```

