---
  title: "CAR Model, Alaska"
author: "Claire Kelling"
date: "June 11, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this code is to fit the CAR model to voting data, aggregated by ward. We only have Census data by district and/or block group so this presents a spatial mismatch file. 

First, we will install all necessary packages.
```{r}
# Packages:
library(sp)
library(spdep)
library(classInt)
library(fields)
library(ggplot2)
library(dplyr)
library(ade4) 
library(igraph) 
library(CARBayesdata)
library(CARBayes)
library(gridExtra)
library(xtable)
library(stringi)
require(rgdal)
library(ngspatial)

```

Now, we will load the data, including the following:
  * shape files
* demographics
* voting data
```{r}
#precinct shape files, showing some precincts to the right of the main figure
prec_shp <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2013-SW-Precinct-Proc-Plan", layer = "2013-SW-Proc-Shape-files")

#dist_shp <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2013-HD-ProclamationPlan", layer = "2013ProclamationPlan")

#voting and demographic data
demog_vote <- read.csv("C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/AKDataByPrecinct.csv")

#redo shape file from github source
# https://github.com/aaron-strauss/precinct-shapefiles/blob/master/ak/2012_ak_precincts.zip
#prec_shp2 <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2012_ak_precincts", layer = "SW Amended Precinct shape files", p4s = "+init=epsg:3338")
```

Exploring shape file
```{r}
plot(prec_shp)
#plot(dist_shp)

#new shape file for precincts
#plot(prec_shp2) #still has the precincts on the right hand side of the graph
```


Now we need to join the spatial polygons and the demographic/voting data.
```{r}
#re-formatting to add the data to the 
View(prec_shp@data)

#check classes
class(prec_shp@data$DISTRICT)

#change column name for ID variables for which we will join
colnames(demog_vote)[1] <- "GEOID"
colnames(prec_shp@data)[3] <- "GEOID"

#change to character variable
demog_vote$GEOID <- as.character(demog_vote$GEOID)
prec_shp@data$GEOID <- as.character(prec_shp@data$GEOID)

#creating new variable
demog_vote$IncumD_Pop <- demog_vote$Incumbent.D*demog_vote$sumTOTALPOP
demog_vote$IncumR_Pop <- demog_vote$Incumbent.R*demog_vote$sumTOTALPOP

#join on GEOID
prec_shp@data <- left_join(prec_shp@data, demog_vote, by = (GEOID = "GEOID"))

length(which(is.na(prec_shp@data$sumRENTED)))
nrow(prec_shp@data)

nrow(demog_vote)
nrow(prec_shp@data)

class(demog_vote$GEOID)
class(prec_shp@data$GEOID)

#we have successfully joined the two datasets
```


Our first step in terms of modeling will be to run our model in a linear model and to test for spatial dependence. 
```{r}
#we need to decide variables to use in our analysis
colnames(demog_vote)
#contested - B if it is contested
#Mayer: D.Pres.Vote, R.Pres.Vote, sumTOTALPOP, sumBLACK, sumHISPANIC
#Mayer Response: D.Dist.Vote, R.Dist.Vote
#TP (third party)

#approximately 50% of the data is contested (vs uncontested)
#right now, I will just use the uncontested data
length(c(which(is.na(prec_shp$D.Dist.Vote)), which(is.na(prec_shp$R.Dist.Vote))))
unc <- unique(c(which(is.na(prec_shp$D.Dist.Vote)), which(is.na(prec_shp$R.Dist.Vote))))

con_shp <- prec_shp[-unc,]

#non-spatial modeling
d.form <- D.Dist.Vote ~ D.Pres.Vote + R.Pres.Vote+sumTOTALPOP +sumBLACK + sumWHITE + sumHISPANIC+ IncumD_Pop + IncumR_Pop
r.form <- R.Dist.Vote ~ D.Pres.Vote + R.Pres.Vote+sumTOTALPOP +sumBLACK + sumWHITE + sumHISPANIC+ IncumD_Pop + IncumR_Pop
#d.model <- lm(formula=d.form, data=prec_shp@data)
#d.resid.model <- residuals(d.model)
#r.model <- lm(formula=r.form, data=prec_shp@data)
#r.resid.model <- residuals(r.model)

c.d.model <- lm(formula=d.form, data=con_shp@data)
c.d.resid.model <- residuals(c.d.model)
c.r.model <- lm(formula=r.form, data=con_shp@data)
c.r.resid.model <- residuals(c.r.model)



#glm model
#d.glm_model <- glm(formula=d.form, family = "poisson", data=prec_shp@data)
#d.resid.glm <- residuals(d.glm_model)
#r.glm_model <- glm(formula=r.form, family = "poisson", data=prec_shp@data)
#r.resid.glm <- residuals(r.glm_model)

c.d.glm_model <- glm(formula=d.form, family = "poisson", data=con_shp@data)
c.d.resid.glm <- residuals(c.d.glm_model)
c.r.glm_model <- glm(formula=r.form, family = "poisson", data=con_shp@data)
c.r.resid.glm <- residuals(c.r.glm_model)


#need to remove shape file objects with no D.Dist.Vote
#no_dem <- prec_shp[-which(is.na(prec_shp$D.Dist.Vote)),]
#no_rep <- prec_shp[-which(is.na(prec_shp$R.Dist.Vote)),]


#test for spatial dependence
#null hypothesis of no spatial autocorrelation (alternative of positive spatial autocorrelation)
#also computes Moran's I statistic 
#if p-value < 0.05, we conclude there is positve spatial autocorrelation
#W.nb <- poly2nb(prec_shp, row.names = rownames(prec_shp@data))
#W.nb.nod <- poly2nb(no_dem, row.names = rownames(no_dem@data))
#W.nb.nor <- poly2nb(no_rep, row.names = rownames(no_rep@data))
W.nb.con <- poly2nb(con_shp, row.names = rownames(con_shp@data))

#coords <- coordinates(prec_shp)
#plot(prec_shp)
#plot(W.nb, coords, pch = 1, cex = 0.6, add = TRUE)

#remove empty neighbor sets

##############
### Creating adjacency matrix from nb
##############
#W_mat <- nb2mat(W.nb, style="B", zero.policy=TRUE) #includes some islands
#View(head(W_mat))
#W_mat_nod <- nb2mat(W.nb.nod, style="B", zero.policy=TRUE)
#W_mat_nor <- nb2mat(W.nb.nor, style="B", zero.policy=TRUE)
W_mat_con <- nb2mat(W.nb.con, style="B", zero.policy=TRUE)


#############
### Moran's I Test
#############

#addition
#W.list <- nb2listw(W.nb, style="B", zero.policy = TRUE)
#W.list.nod <- nb2listw(W.nb.nod, style="B", zero.policy = TRUE)
#W.list.nor <- nb2listw(W.nb.nor, style="B", zero.policy = TRUE)
W.list.con <- nb2listw(W.nb.con, style="B", zero.policy = TRUE)

#moran.mc(x=resid.glm, listw=W.list, nsim=1000, zero.policy = FALSE)
#moran.mc(x=d.resid.glm, listw=W.list.nod, nsim=1000, zero.policy = TRUE)
#moran.mc(x=r.resid.glm, listw=W.list.nor, nsim=1000, zero.policy = TRUE)

moran.mc(x=c.d.resid.glm, listw=W.list.con, nsim=1000, zero.policy = TRUE)
moran.mc(x=c.r.resid.glm, listw=W.list.con, nsim=1000, zero.policy = TRUE)

#if p-value < 0.05, we conclude there is positve spatial autocorrelation

```


Need to eliminate the islands with no neighbors
```{r}
no_nb <- which(rowSums(W_mat_con) == 0)


nonb_W <- W_mat_con[-no_nb, -no_nb]
nonb_shp <- con_shp[-no_nb,]

nonb_Wnb <- poly2nb(nonb_shp, row.names = rownames(nonb_shp@data))
```


Now, we will fit both the Leroux and the BYM model.
```{r}
##############
#Leroux Model
##############
#d.model.ler <- S.CARleroux(formula=d.form, data=no_dem@data,
#                             family="poisson", W=W_mat_nod, burnin=20000, n.sample=120000, thin=10)

#r.model.ler <- S.CARleroux(formula=r.form, data=no_rep@data,
#                             family="poisson", W=W_mat_nor, burnin=20000, n.sample=120000, thin=10)


c.d.model.ler <- S.CARleroux(formula=d.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)

c.r.model.ler <- S.CARleroux(formula=r.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)

##############
# BYM Model
##############
#d.model.bym <- S.CARbym(formula=d.form, data=no_dem@data,
#                          family="poisson", W=W_mat_nod, burnin=20000, n.sample=120000, thin=10)

#r.model.bym <- S.CARbym(formula=r.form, data=no_rep@data,
#                          family="poisson", W=W_mat_nor, burnin=20000, n.sample=120000, thin=10)

c.d.model.bym <- S.CARbym(formula=d.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)


c.r.model.bym <- S.CARbym(formula=r.form, data=nonb_shp@data,
                             family="poisson", W=nonb_W, burnin=20000, n.sample=120000, thin=10)

##############
# Model Evaluation
##############
#  dem ler
#d.model.ler$modelfit
#  rep ler
#d.model.ler$modelfit
#  dem bym
#r.model.bym$modelfit
#  rep bym
#r.model.bym$modelfit

#Only using contested precincts:
#  dem ler
c.d.model.ler$modelfit
#  rep ler
c.r.model.ler$modelfit
#  dem bym
c.d.model.bym$modelfit
#  rep bym
c.r.model.bym$modelfit
```

Now, I will also include a model that accounts for spatial autocorrelation.
```{r}
#d.sp.sglmm.fit <- sparse.sglmm(formula = d.form,data=no_dem@data, family = poisson, A = W_mat_nod,
#                             verbose = TRUE) #tune = list(sigma.s = 0.02)

#r.sp.sglmm.fit <- sparse.sglmm(formula = r.form,data=no_rep@data, family = poisson, A = W_mat_nor,
#                             verbose = TRUE) #tune = list(sigma.s = 0.02)

rownames(nonb_W) <- NULL
c.d.sp.sglmm.fit <- sparse.sglmm(formula = d.form,data=nonb_shp@data, family = poisson, 
                                 A = nonb_W, verbose = TRUE) #tune = list(sigma.s = 0.02)

#25 min at 9:43, stuck at 44%, 23m 49s

c.r.sp.sglmm.fit <- sparse.sglmm(formula = r.form,data=nonb_shp@data, family = poisson, 
                                 A = nonb_W, verbose = TRUE) #tune = list(sigma.s = 0.02)
#11%, 32m 22s remaining at 9:56pm
#55% ~19m 27s  

#save the models for later reference, takes over an hour to run
save(c.d.model.ler, c.r.model.ler, c.d.model.bym, c.r.model.bym, 
     #c.d.sp.sglmm.fit, c.r.sp.sglmm.fit, 
     file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/
     Alaska Project/prelim_fit/prelim_mod_fit.Rdata")

#summary(sp.sglmm.fit)
#d.sp.sglmm.fit$dic
#d.sp.sglmm.fit$pD

#r.sp.sglmm.fit$dic
#r.sp.sglmm.fit$pD

c.d.sp.sglmm.fit$dic
c.d.sp.sglmm.fit$pD

c.r.sp.sglmm.fit$dic
c.r.sp.sglmm.fit$pD

```

Now, we need to compare coefficients to see if there was any spatial autocorrelation.
```{r}
model.bym$summary.results[,1:3]

summary(c.d.sp.sglmm.fit)
```

#Model Evaluation
Now, I will compare the models using mean-squared prediction error.

First, I need to set up predicted for each model and actual for both Democratic and Republican races.
```{r}
actual.d <- nonb_shp@data$D.Dist.Vote
actual.r <- nonb_shp@data$R.Dist.Vote

d.pred.bym <- c.d.model.bym$fitted.values
r.pred.bym <- c.r.model.bym$fitted.values

d.pred.ler <- c.d.model.ler$fitted.values
r.pred.ler <- c.r.model.ler$fitted.values

#d.pred.sglmm <- c.d.sp.sglmm.fit$fitted.values
#r.pred.sglmm <- c.r.sp.sglmm.fit$fitted.values

```

Now, I will compute the MSPE for each model.
```{r}
#### Democratic Prediction:
#linear model
d.lm.mspe <- mean((c.d.resid.model)^2)
#glm
d.glm.mspe <- mean((c.d.resid.glm)^2)
#bym
d.bym.mspe <- mean((d.pred.bym-actual.d)^2)
#leroux
d.ler.mspe <- mean((d.pred.ler-actual.d)^2)
#sglmm
#d.sglmm.mspe <- mean((d.pred.sglmm - actual.d)^2)

#### Republican Prediction:
#linear model
r.lm.mspe <- mean((c.r.resid.model)^2)
#glm
r.glm.mspe <- mean((c.r.resid.glm)^2)
#bym
r.bym.mspe <- mean((r.pred.bym-actual.r)^2)
#leroux
r.ler.mspe <- mean((r.pred.ler-actual.r)^2)
#sglmm
#r.sglmm.mspe <- mean((r.pred.sglmm - actual.r)^2)

mspe_tab <- rbind(c(d.lm.mspe, d.glm.mspe, d.bym.mspe, d.ler.mspe),#, d.sglmm.mspe),
                  c(r.lm.mspe, r.glm.mspe, r.bym.mspe, r.ler.mspe))#, r.sglmm.mspe))
colnames(mspe_tab) <- c("lin", "glm", "bym", "leroux")#, "sglmm")
rownames(mspe_tab) <- c("Democrat", "Republican")

```


Bayesian Model Comparison Techniques:
```{r}
# dem ler
c.d.model.ler$modelfit
#  rep ler
c.r.model.ler$modelfit
#  dem bym
c.d.model.bym$modelfit
#  rep bym
c.r.model.bym$modelfit

#c.d.sp.sglmm.fit$dic
#c.d.sp.sglmm.fit$pD

#c.r.sp.sglmm.fit$dic
#c.r.sp.sglmm.fit$pD
```


Creation of counterfactual scenario where there is no incumbent:
```{r}
#First, I will create the model dataframe
mod_dat <- demog_vote[,c("D.Pres.Vote", "R.Pres.Vote", "sumTOTALPOP", "sumBLACK", "sumWHITE", "sumHISPANIC", "IncumD_Pop", "IncumR_Pop")]
mod_dat$int <- rep(1, nrow(mod_dat))
mod_dat <- cbind(mod_dat[,9], mod_dat[,1:8])


#create function that returns predicted values under counterfactual situation
pred_val <- function(coeffs.Dem, coeffs.Rep){
  coeffs.Dem[8] <- 0
  coeffs.Dem[9] <- 0
  coeffs.Rep[8] <- 0
  coeffs.Rep[9] <- 0
  
  tableDem <- as.matrix(mod_dat)%*%coeffs.Dem
  tableDem <- as.numeric(tableDem)
  tableDem <- cbind(tableDem, demog_vote$District)
  tableDem <- as.data.frame(tableDem)
  colnames(tableDem) <- c("votes", "District")
  
  length(unique(tableDem$District))
  
  v_table_d <- tableDem %>% group_by(District) %>% dplyr::summarise(dem_votes=sum(votes))
  
  tableRep <- as.matrix(mod_dat)%*%coeffs.Rep
  tableRep <- as.numeric(tableRep)
  tableRep <- cbind(tableRep, demog_vote$District)
  tableRep <- as.data.frame(tableRep)
  colnames(tableRep) <- c("votes", "District")
  
  v_table_r <- tableRep %>% group_by(District) %>% dplyr::summarise(rep_votes=sum(votes))
  
  v_table <- left_join(v_table_d, v_table_r)
  
  colnames(v_table) <- c("District", "Dem. Votes", "Rep. Votes")
  return(v_table)
}


######
#Linear Model:
######

coeffs.Dem <-summary(c.d.model)$coefficients[,1]
coeffs.Rep <-summary(c.r.model)$coefficients[,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_lm <- pred_val(coeffs.Dem, coeffs.Rep)

######
#GLM:
######

coeffs.Dem <-summary(c.d.glm_model)$coefficients[,1]
coeffs.Rep <-summary(c.r.glm_model)$coefficients[,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_glm <- pred_val(coeffs.Dem, coeffs.Rep)

######
#BYM:
######

coeffs.Dem <- c.d.model.bym$summary.results[1:9,1]
coeffs.Rep <- c.r.model.bym$summary.results[1:9,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_bym <- pred_val(coeffs.Dem, coeffs.Rep)

######
#Leroux:
######

coeffs.Dem <- c.d.model.ler$summary.results[1:9,1]
coeffs.Rep <- c.r.model.ler$summary.results[1:9,1]

coeffs.Dem <- as.numeric(coeffs.Dem)
coeffs.Rep <- as.numeric(coeffs.Rep)

v_table_ler <- pred_val(coeffs.Dem, coeffs.Rep)

######
#SGLMM:
######

# coeffs.Dem <- c.d.sp.sglmm.fit$coefficients
# coeffs.Rep <- c.r.sp.sglmm.fit$coefficients
# 
# coeffs.Dem <- as.numeric(coeffs.Dem)
# coeffs.Rep <- as.numeric(coeffs.Rep)
# 
# v_table_sglmm <- pred_val(coeffs.Dem, coeffs.Rep)

write.csv(v_table_lm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/mod_fit/lin_mod.csv')
write.csv(v_table_glm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/mod_fit/glm.csv')
write.csv(v_table_bym, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/mod_fit/bym.csv')
write.csv(v_table_ler, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/mod_fit/ler.csv')
#write.csv(v_table_sglmm, file = 'C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/mod_fit/sglmm.csv')```
```


To Do: 
* need to also incorporate incumbency
* cross validation statistics?
* try including and not the county indicator
* hypotheticaly if there were no incumbents, fit the model on precincts and then aggregate to the house district level, and see who would win
** in other words, we want two columns (repub votes and dem votes) for each of the 40 districts - do a group_by
* Histogram EG and # D Seats