---
  title: "Heat Maps, Alaska"
author: "Claire Kelling"
date: "June 14, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this code is to create heat maps for Alaska.

First, we will install all necessary packages.
```{r}
# Packages:
library(sp)
library(spdep)
library(classInt)
library(fields)
library(ggplot2)
library(dplyr)
library(ade4) 
library(igraph) 
library(CARBayesdata)
library(CARBayes)
library(gridExtra)
library(xtable)
library(stringi)
require(rgdal)
library(ngspatial)
library(plyr)
library(ggmap) 
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(raster)
library(utils)
library(rgdal)

```

Now, we will load the data, including the following:
  * shape files
* demographics
* voting data
```{r}
#precinct shape files, showing some precincts to the right of the main figure
prec_shp <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2013-SW-Precinct-Proc-Plan", layer = "2013-SW-Proc-Shape-files")

dist_shp <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2013-HD-ProclamationPlan", layer = "2013ProclamationPlan")

#voting and demographic data
#demog_vote <- read.csv("C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/AKDataByPrecinct.csv")

#redo shape file from github source
# https://github.com/aaron-strauss/precinct-shapefiles/blob/master/ak/2012_ak_precincts.zip
#prec_shp2 <- readOGR(dsn = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/2012_ak_precincts", layer = "SW Amended Precinct shape files", p4s = "+init=epsg:3338")
```


Create heatmap based off of the % demographic
```{r}
#take out shape that is out to the right
prec_shp <- prec_shp[-324,]

heat_dat <- read.csv("C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/AKDataByPrecinct(1).csv")

ball_dat <- read.csv("C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/ballot_data.csv")

heat_dat <- heat_dat[-nrow(heat_dat),] #last row is NA

#change column name for ID variables for which we will join
colnames(heat_dat)[1] <- "GEOID"
colnames(prec_shp@data)[3] <- "GEOID"
colnames(ball_dat)[1] <- "GEOID"

#change to character variable
heat_dat$GEOID <- as.character(heat_dat$GEOID)
prec_shp@data$GEOID <- as.character(prec_shp@data$GEOID)
ball_dat$GEOID <- as.character(ball_dat$GEOID)

prec_shp@data$id <- rownames(prec_shp@data)
prec_shp@data <- left_join(prec_shp@data, heat_dat, by = (GEOID = "GEOID"))
prec_shp@data <- left_join(prec_shp@data, ball_dat, by = (GEOID = "GEOID"))

write.csv(prec_shp@data, "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/comb_data.csv")

P4S.latlon <- CRS("+proj=longlat +datum=WGS84")

dist_shp <- spTransform(dist_shp, CRS("+init=epsg:3338")) #4326 or 3338
prec_shp <- spTransform(prec_shp, CRS("+init=epsg:3338"))

## Plot for Proportion
#prec_shp@data   <- join(prec_shp@data, data, by="CD_GEOCODI")
sp_f     <- fortify(prec_shp)
sp_f     <- join(sp_f,prec_shp@data, by="id")
sp_f$Percent_DVote <- as.numeric(as.character(sp_f$Percent_DVote))
#sp_f <- sp_f[which(sp_f$long < 1e10),]

dist_shp_f <- fortify(dist_shp)

fit_by_prec <- ggplot()+ 
  #geom_polygon(data=dist_shp,aes(long,lat, group = group), fill = NA, col = "red") + 
  geom_polygon(data = sp_f, aes(long, lat, group = group, fill = Percent_DVote)) + 
  coord_equal() +
  labs(fill = "Percent Dem Vote")+ 
  geom_polygon(data=dist_shp_f,aes(long,lat, group = group), fill = NA, col = "red") +
  ggtitle("Percentage of Dem Vote by Precinct")+ scale_fill_gradient2(midpoint = 0.5,low = "red", high = "navyblue")
#+theme(text = element_text(size=30))+theme(axis.text.x=element_text(size=20))


## Plot for Proportion
prop_by_prec <- ggplot()+ 
  #geom_polygon(data=dist_shp,aes(long,lat, group = group), fill = NA, col = "red") + 
  geom_polygon(data = sp_f, aes(long, lat, group = group, fill = (Reg..D.Voters-Reg..R.Voters)/Reg..Voters)) + 
  coord_equal() +
  labs(fill = "Reg Dem - Reg Rep / Total Reg")+ 
  geom_polygon(data=dist_shp_f,aes(long,lat, group = group), fill = NA, col = "red") +
  ggtitle("Reg Dem - Reg Rep / Total Reg by Precinct")+
  scale_fill_gradient(low = "lightblue", high = "navyblue")


sp_f     <- fortify(prec_shp)
sp_f     <- join(sp_f,prec_shp@data, by="id")
sp_f$Percent_DVote <- as.numeric(as.character(sp_f$Percent_DVote))
sp_f <- sp_f[which(sp_f$long < 1e10),]
dist_shp_f <- fortify(dist_shp)


#Plot for ballot issue
ballot_by_prec <- ggplot()+ 
  #geom_polygon(data=dist_shp,aes(long,lat, group = group), fill = NA, col = "red") + 
  geom_polygon(data = sp_f, aes(long, lat, group = group, fill = (SJR2.Yes)/(SJR2.Yes + SJR2.No))) + 
  coord_equal() +
  labs(fill = "Yes on SJR2 as Percentage of Total")+ 
  geom_polygon(data=dist_shp_f,aes(long,lat, group = group), fill = NA, col = "red") +
  ggtitle("Percentage of Total")+ 
  scale_fill_gradient2(midpoint = 0.5,low = "red", high = "navyblue")

```

Make the plot into a function
```{r}

plot_fun <- function(dat, var_name, mid, title){
  plot_by_prec <- ggplot()+ 
    #geom_polygon(data=dist_shp,aes(long,lat, group = group), fill = NA, col = "red") + 
    geom_polygon(data = dat, 
                 aes(long, lat, group = group, fill = var_name)) + 
    coord_equal() +
    labs(fill = title)+ 
    geom_polygon(data=dist_shp_f,aes(long,lat, group = group), fill = NA, col = "red") +
    ggtitle(title)+ 
    scale_fill_gradient2(midpoint = mid,low = "red", high = "navyblue")
}

```


Now, we want to zoom in to Anchorage
```{r}
#these are the districts for Anchorage, possibly including 16:28
anch_dist <- 16:26

#length(unique(sp_f$District))
#head(unique(sp_f$District))
#class(sp_f$District)

sp_anch <- sp_f[which(sp_f$District %in% anch_dist),]
dist_shp_f$dist <- as.numeric(dist_shp_f$id) +1
dist_shp_anch <- dist_shp_f[which(dist_shp_f$dist %in% anch_dist),]

#unique(dist_shp_f$id)
#unique(sp_f$District)

fit_by_anch <- ggplot()+ 
  geom_polygon(data = sp_anch, aes(long, lat, group = group, fill = Percent_DVote)) + 
  coord_equal() +
  labs(fill = "Percent Dem Vote")+ 
  geom_polygon(data=dist_shp_anch,aes(long,lat, group = group), fill = NA, col = "red") +
  ggtitle("Percentage of Dem Vote by Precinct")+ scale_fill_gradient2(midpoint = 0.5,low = "red", high = "navyblue")
```

Plotting district Alignments
```{r}
dist_dat <- read.csv("C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/data/district_alignments.csv")
colnames(dist_dat)[1] <- "dist"

dist_shp_f <- left_join(dist_shp_f, dist_dat)


fit_dist <- ggplot()+ 
  #geom_polygon(data=dist_shp,aes(long,lat, group = group), fill = NA, col = "red") + 
  geom_polygon(data = dist_shp_f, aes(long, lat, group = group, fill = Prob..Dem.)) + 
  coord_equal() +
  labs(fill = "Prob Dem")+ 
  geom_polygon(data=dist_shp_f,aes(long,lat, group = group), fill = NA, col = "white") +
  ggtitle("Prob Dem")+ scale_fill_gradient2(midpoint = 0.5,low = "red", high = "navyblue")

fit_dist_anch <- ggplot()+ 
  #geom_polygon(data=dist_shp,aes(long,lat, group = group), fill = NA, col = "red") + 
  geom_polygon(data = dist_shp_f[which(dist_shp_f$dist %in% anch_dist),], aes(long, lat, group = group, fill = Prob..Dem.)) + 
  coord_equal() +
  labs(fill = "Prob Dem")+ 
  geom_polygon(data=dist_shp_f[which(dist_shp_f$dist %in% anch_dist),],aes(long,lat, group = group), fill = NA, col = "white") +
  ggtitle("Prob Dem")+ scale_fill_gradient2(midpoint = 0.5,low = "red", high = "navyblue")

range(dist_shp_f$Prob..Dem.)
```

Starting with attempt to add Google Maps in the background
```{r}
gmap <- get_map(location=b)
mgmap <- as.matrix(gmap)
vgmap <- as.vector(mgmap)
vgmaprgb <- col2rgb(vgmap)
gmapr <- matrix(vgmaprgb[1, ], ncol = ncol(mgmap), nrow = nrow(mgmap))
gmapg <- matrix(vgmaprgb[2, ], ncol = ncol(mgmap), nrow = nrow(mgmap))
gmapb <- matrix(vgmaprgb[3, ], ncol = ncol(mgmap), nrow = nrow(mgmap))
rgmaprgb <- brick(raster(gmapr), raster(gmapg), raster(gmapb))
rm(gmapr, gmapg, gmapb)
extent(rgmaprgb) <- unlist(attr(gmap,which="bb"))[c(2,4,1,3)]
projection(rgmaprgb) <- CRS("+init=epsg:3338")

plot(rgmaprgb)

rprobextSpDF <- as(extent(unlist(attr(gmap))), 
    "SpatialPolygons")
projection(rprobextSpDF) <- CRS("+init=epsg:4326")
rprobextGM <- spTransform(rprobextSpDF, CRS("+init=epsg:3338"))

library(sp)
library(rworldmap)

b <- bbox(prec_shp)
alaska_base <- ggmap(get_map(location = b))

alaska_base+ 
  #geom_polygon(data=dist_shp,aes(long,lat, group = group), fill = NA, col = "red") + 
  geom_polygon(data = sp_f, aes(long, lat, group = group, fill = Percent_DVote))  +
  labs(fill = "Percent Dem Vote")+ 
  geom_polygon(data=dist_shp_f,aes(long,lat, group = group), fill = NA, col = "red") +
  ggtitle("Percentage of Dem Vote by Precinct")+ scale_fill_gradient2(midpoint = 0.5,low = "red", high = "navyblue")+ 
  coord_equal()#+
  #coord_map("cylindrical")

library(ggalt)

```

