---
  title: "CAR Model, Alaska"
author: "Claire Kelling"
date: "June 11, 2018"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this code is to fit the CAR model to voting data, aggregated by ward. We only have Census data by district and/or block group so this presents a spatial mismatch file. 

First, we will install all necessary packages.
```{r}
# Packages:
library(sp)
library(spdep)
library(classInt)
library(fields)
library(ggplot2)
library(dplyr)
library(ade4) 
library(igraph) 
library(CARBayesdata)
library(CARBayes)
library(gridExtra)
library(xtable)
library(stringi)
```

Now, we will load the data, including the following:
  * shape files
* demographics
* voting data
```{r}
#shape files
load(file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/precinct_shape.Rdata")
#demographics
load(file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/demog.Rdata")
#voting data
load(file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/vote.Rdata")
```

Now we need to join the spatial polygons and the demographic/voting data.
```{r}
#re-formatting to add the data to the SpatialPolygonsDataFrame
ward_shape$id <- row.names(ward_shape)
ward_shape@data <- left_join(ward_shape@data, demog, by = (GEOID = "GEOID"))
ward_shape@data <- left_join(ward_shape@data, vote, by = (GEOID = "GEOID"))
```


Our first step in terms of modeling will be to run our model in a linear model and to test for spatial dependence. 
```{r}
#non-spatial modeling
form <- vote_count ~ median_income + upemp_rate+total_pop+perc_male+med_age+herf_index
model <- lm(formula=form, data=ward_shape@data)
resid.model <- residuals(model)

#test for spatial dependence
#null hypothesis of no spatial autocorrelation (alternative of positive spatial autocorrelation)
#also computes Moran's I statistic 
#if p-value < 0.05, we conclude there is positve spatial autocorrelation
W.nb <- poly2nb(ward_shape, row.names = rownames(ward_shape@data))

##############
### Creating adjacency matrix from nb
##############
W_mat <- nb2mat(W.nb, style="B")

#############
### Moran's I Test
#############

#addition
W.add.list <- nb2listw(add_nb, style="B")
moran.mc(x=resid.model, listw=W.add.list, nsim=1000)

```


Now, we will fit both the Leroux and the BYM model.
```{r}
##############
#Leroux Model
##############
#binary
model.ler.bin <- S.CARleroux(formula=form, data=ward_shape@data,
                             family="poisson", W=bin_W, burnin=20000, n.sample=120000, thin=10)

##############
# BYM Model
##############
#binary
model.bym.bin <- S.CARbym(formula=form, data=ward_shape@data,
                          family="poisson", W=add_W, burnin=20000, n.sample=120000, thin=10)

#save the models for later reference, takes over an hour to run
save(model.bym.add, model.bym.bin, model.ler.add, model.ler.bin, file = "C:/Users/ckell/Dropbox/Gerrycamp Regressions Project/Alaska Project/prelim_fit/ward_shape.Rdata")

##############
# BYM Model
##############
#  add bym
model.bym.add$modelfit
#  bin bym
model.bym.bin$modelfit
#  add ler
model.ler.add$modelfit
#  bin ler
model.ler.bin$modelfit
```

Now, I will also include a model that accounts for spatial autocorrelation.
```{r}
sp.sglmm.fit <- sparse.sglmm(formula = form,data=ward_shape_geog@data, family = poisson, A = W_mat,
                             verbose = TRUE) #tune = list(sigma.s = 0.02)


#summary(sp.sglmm.fit)
sp.sglmm.fit$dic
sp.sglmm.fit$pD
```

Now, we need to compare coefficients to see if there was any spatial autocorrelation.
```{r}
model.bym.geog$summary.results[,1:3]

summary(sp.sglmm.fit.soc)
```

